### 2020-10-01 reports/article.tex
simplifying - no calibration

\documentclass[10pt,letterpaper]{article}
\usepackage[top=0.85in,left=2.75in,footskip=0.75in,marginparwidth=2in]{geometry}

% use Unicode characters - try changing the option if you run into troubles with special characters (e.g. umlauts)
\usepackage[utf8]{inputenc}

% clean citations
\usepackage{cite}

% hyperref makes references clicky. use \url{www.example.com} or \href{www.example.com}{description} to add a clicky url
\usepackage{nameref,hyperref}

% line numbers
\usepackage[right]{lineno}

% improves typesetting in LaTeX
\usepackage{microtype}
\DisableLigatures[f]{encoding = *, family = * }

% text layout - change as needed
\raggedright
\setlength{\parindent}{0.5cm}
\textwidth 5.25in 
\textheight 8.75in

% Remove % for double line spacing
%\usepackage{setspace} 
%\doublespacing

% use adjustwidth environment to exceed text width (see examples in text)
\usepackage{changepage}

% adjust caption style
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,singlelinecheck=off]{caption}

% remove brackets from references
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% headrule, footrule and page numbers
\usepackage{lastpage,fancyhdr,graphicx}
\usepackage{epstopdf}
\pagestyle{myheadings}
\pagestyle{fancy}
\fancyhf{}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
\fancyheadoffset[L]{2.25in}
\fancyfootoffset[L]{2.25in}

% use \textcolor{color}{text} for colored text (e.g. highlight to-do areas)
\usepackage{color}

% define custom colors (this one is for figure captions)
\definecolor{Gray}{gray}{.25}

% this is required to include graphics
\usepackage{graphicx}

% use if you want to put caption to the side of the figure - see example in text
\usepackage{sidecap}

% use for have text wrap around figures
\usepackage{wrapfig}
\usepackage[pscoord]{eso-pic}
\usepackage[fulladjust]{marginnote}
\reversemarginpar

% for better-looking tables
\usepackage{booktabs}

% math utils
\usepackage{amsmath}

% symbols
\usepackage{gensymb}

% % citation aliases
% \usepackage{natbib}
% % https://tex.stackexchange.com/questions/323868/setting-two-different-aliases-name-2016-name-2016-with-defcitealias
% \newcommand{\mycitet}[1]{\citetalias{#1} (\citeyear{#1})}
% \newcommand{\mycitep}[1]{(\citetalias{#1} \citeyear{#1})}

% force floats in sections
\usepackage{float}

% \usepackage{caption}

% document begins here
\begin{document}
\vspace*{0.35in}

% \defcitealias{arezh2018bodenbedeckung}{ARE}
% \defcitealias{asitvd2018structure}{ASIT VD}
% \defcitealias{bve2019mopube}{BVE}

% title goes here:
\begin{flushleft}
{\Large Spatially-explicit simulation of urban heat mitigation by increasing the tree canopy cover
\textbf\newline{}
}
\newline
% authors go here:
\\
Mart\'i Bosch\textsuperscript{*},
% Maxence Locatelli,
% J\'er\^ome Chenal
\\
\bigskip
Urban and Regional Planning Community (CEAT), \'Ecole Polytechnique F\'ed\'erale de Lausanne (EPFL), Lausanne, Switzerland
\\ 
\bigskip
* Corresponding author: marti.bosch@epfl.ch

\end{flushleft}

\section*{Abstract}

% now start line numbers
\linenumbers

% the * after section prevents numbering
\section*{Introduction}

% Distinction: three types of UHI: urban boundary layer, urban canopy layer \cite{oke1976distinction}

% ``We know that simply increasing the percent cover of greenspace leads to a reduction of temperatures; this relationship is very consistent. What is less known, however, is the effects of the spatial configuration of that greenspace on urban temperatures'' \cite[page 2]{zhou2017effects}
% The complexity of such relationship might only be understood by means of a spatially-explicit analysis. Several studies have addressed the effects of landscape composition and pattern on land surface temperature (LST), e.g., derived from satellite imagery, by means of statistical approaches (e.g., ordinary regressions or spatial autoregressions). 
% non-linear interplay between landscape patterns and ecological processes \cite{du2016quantifying}
% Note that while most studies that use single-level statistical models suggest that landscape composition has greater influence on LST than landscape configuration, the multi-level models used by Du et al. \cite{du2016quantifying} suggest the opposite. Why? (i) group-based configuration metrics consider the autocorrelations of the configuration metrics, and (ii) the aggregation effects of the groups and the individual effects of the pixels are distinguished by the multi-level models.
% Note however further limitations of statistical approaches in general:
% ``Our results from the comparison of the two cities indicated that the spatial configuration of trees may have different effects on LST in cities with different climatic conditions'' \cite[page 6]{zhou2017effects}
% This might be due to the fact that the relative contributions of the two main cooling mechanisms associated to urban trees, namely shading and evapotranspiration, can be related to the different climatic conditions of the modelled cities \cite{zhou2017effects}.
% Nevertheless, the statistical approaches employed in such studies (e.g., ordinary regressions or spatial autoregressions with LST as response and landscape metrics as explanatory variables) do not explicitly reflect the physical mechanisms beyond the urban heat island phenomena.

% MAYBE: greening \cite{jabareen2006sustainable}

Urbanization is nowadays a global phenomenon that is increasingly concentrating the world's population in urban areas, which are expected to grow in both the number of dwellers and spatial extent \cite{seto2011meta, angel2012atlas, unitednations2018world}.
As a major force of landscape change, urbanization is characterized by the conversion of natural to artificial surfaces, which alters the energy and water exchanges as well as the movement of air. Such changes often result in the urban heat island (UHI) effect, a phenomena by which urban temperatures are warmer than its rural surroundings \cite{oke1982energetic}.

% As urban areas grow and global temperatures increase, heat waves are expected to become more intense and frequent ant the UHI effect
As urban areas grow and global temperatures rise, the UHI effect is expected to become more intense \cite{meehl2004more,huang2019projecting}, which makes urban heat mitigation a major priority for urban planning and policy-making.
% Green infrastructure
% To that end, cities are increasingly
Increasing urban green space, especially the urban tree canopy, has been one of the most widely advocated strategies of urban heat mitigation \cite{bowler2010urban,phelan2015urban,koc2018evaluating}.
% Overall decrease of urban green space in European cities due to infill development \cite{haaland2015challenges}.
% TODO: challanges with densification, trade-offs in the provision of ecosystem services
% TODO: need for methods to quantify
Nevertheless, the impact of the abundance and spatial configuration of the urban tree canopy on air temperature shows a complex spatial behaviour that remains
Therefore, there is an urgent need to develop robust and spatially explicit methods to quantify the contribution of urban green space to urban heat mitigation \cite{haaland2015challenges,artmann2019urban}.

The present work introduces a novel approach to spatially evaluate the heat mitigation potential that could be achieved by increasing the urban tree canopy cover in the urban agglomeration of Lausanne, Switzerland. The proposed method consists of three major parts.
% 1. Couple the cadastral LULC map with a high-resolution tree canopy map to detect locations where the tree canopy cover can be increased
% 2. Generate a variety of synthetic scenarios by increasing the tree canopy cover, which can mimic urban greening strategies. % in different spatial configurations
% 3. 
First, the cadastral LULC map is coupled with a high-resolution tree canopy map to detect candidate pixels where the tree canopy cover can be increased.
Secondly, a set of synthetic scenarios is generated by increasing the tree canopy cover in the candidate pixels.
% is applied with the LULC map that corresponds to 
% Finally, the synthetic LULC map of each scenario is fed to the InVEST urban cooling model in order to estimate the air temperature of
Finally, the spatial distribution of air temperature of each scenario is estimated by applying the InVEST urban cooling model to the corresponding synthetic LULC map.

\section*{Materials and Methods}

\subsection*{Study area}

Lausanne is the fourth largest Swiss urban agglomeration with 420757 inhabitants as of January 2019 \cite{sfso2018city}. The agglomeration is situated at the Swiss Plateau and on the shore of the Lake L\'eman, and is characterized by a continental temperate climate with mean annual temperatures of 10.9 $\degree$C and mean annual precipitation of 100 mm, with a dominating vegetation of mixed broadleaf forest. The spatial extent of the study has been selected following the recent application of the InVEST urban cooling model to Lausanne by Bosch et al. \cite{bosch2020spatially}, and covers an area of 112.46 $km^2$.

In order to evaluate the human thermal comfort, the population data for the study area has been extracted from the population and households statistics (STATPOP) \cite{sfso2020statistique} provided by the Swiss Federal Statistical Office (SFSO) with the Python library swisslandstats-geopy \cite{bosch2019swisslandstats}.

\subsection*{Simulation with the InVEST urban cooling model}

The spatial distribution of UHI is simulated with the InVEST urban cooling model (version 3.8.0) \cite{sharp2020invest}, which is based on the heat mitigation provided by shade, evapotranspiration and albedo. The main inputs are a LULC raster map, a reference evapotranspiration raster and a biophysical table containing model information of each LULC class of the map. Each row of the biophysical table represents a LULC class, and features the following columns:

\begin{itemize}
\item \texttt{lucode} the LULC class code as represented in the LULC raster map
\item \texttt{Shade} a value between 0 and 1 representing the proportion of tree cover in such LULC class
\item \texttt{Kc} the evapotranspiration coefficient
\item \texttt{Albedo} a value between 0 and 1 representing the proportion of solar radiation directly reflected by the LULC class
\item \texttt{Green\_area} whether the LULC class should be considered a green area
% \item \texttt{Building\_intensity} a value between 0 and 1 representing the ratio between floor area and land area
\end{itemize}

In order to simulate the spatial distribution of $T_{air}$, the model requires two additional inputs. The first is the rural reference temperature $T_{ref}$, where the UHI effect is not observed, e.g., in the rural surroundings of the city. The second is the magnitude of the urban heat island effect $UHI_{max}$, namely the difference between the rural refrence temperature and the maximum $T_{air}$ observed in the city center.

The data inputs described above are then used to compute the cooling capacity index, which is based on the physical mechanisms that contribute to cooling urban temperatures. More precisely, the cooling capacity index used in InVEST urban cooling model builds upon the indices proposed by Zardo et al. \cite{zardo2017estimating}, which are based on shading and evapotranspiration, and extends them by adding a factor to account for the albedo.
For each pixel $i$ of the LULC raster map, the cooling capacity index is computed as in:

\begin{equation}
  \label{eq:cooling-capacity}
  CC_i = w_{S} \cdot \textrm{S}_i + w_{AL} \cdot \textrm{AL}_i + w_{ET} \cdot \textrm{ETI}_i
\end{equation}

where $S_i$, $AL_i$ and $ETI_i$ respectively represent the tree shading, albedo and evapotranspiration values of pixel $i$ as defined in the biophysical table, and $w_{S}$, $w_{AL}$ and $w_{ET}$ represent the weights attributed to each component respectively.
The values of $S_i$ and $AL_i$ are retrieved from the biophysical table according to the LULC class of the pixel $i$ (see \autoref{tab:biophysical-table}), whereas the reference evapotranspiration $ETI_i$ pixel values are estimated with the Hargreaves equation \cite{hargreaves1985reference} based on the daily minimum, average and maximum $T_{air}$ values of the 1~km gridded inventory of by the Federal Office of Meteorology and Climatology (MeteoSwiss) \cite{frei2014interpolation} (see Bosch et al. \cite{bosch2020spatially} for more details on the estimation of $ETI_i$).
% The values of $S_i$ and $AL_i$ are retrieved from the biophysical table according to the post-refinement LULC class of the pixel $i$, e.g., the shade coefficients are computed as the midpoint of the bin interval of each level of tree cover (i.e., 0.125, 0.375, 0.625 and 0.875), whereas the albedo coefficients have been linearly interpolated based on the level of building cover (see \ref{sec:biophysical-table}). On the other hand, the reference evapotranspiration $ETI_i$ pixel values are estimated with the Hargreaves equation \cite{hargreaves1985reference} based on the daily minimum, average and maximum $T_{air}$ values of the 1~km gridded inventory of by the Federal Office of Meteorology and Climatology (MeteoSwiss) \cite{frei2014interpolation} (see Bosch et al. \cite{bosch2020spatially} for more details on the estimation of $ETI_i$).

In order to account for the cooling effect of large green spaces, the computed cooling capacity index of pixels that are part of large green areas ($>$ 2~ha) is adjusted as in:

\begin{equation}
  \label{eq:cooling-capacity-green}
  CC_i^{green} = \sum_{j \in \Omega_i} g_i \cdot CC_j \cdot e^{-\frac{d(i, j)}{d_{cool}}}
\end{equation}

where $g_i$ is 1 when the pixel $i$ is a green area and 0 otherwise (as defined in the biophyisical table), $d(i, j)$ is the distance between pixels $i$ and $j$, $d_{cool}$ is a parameter that defines the distance over which a green space has a cooling effect, and $\Omega_i$ is the set of pixels whose distance to $i$ is lower than $d_{cool}$.

Then, a heat mitigation index is computed as:

% $\; if \;\; i \; \textrm{is part of a large green area} \;\; or \;\; CC_i > CC_i^{green}$
% $\; otherwise$
\begin{equation}
  \label{eq:heat-mitigation index}
  HM_i=\begin{cases}
    CC_i & if \; i \in \textrm{large green area} \;\; or \;\; CC_i > CC_i^{green} \\
    CC_i^{green} & otherwise \\
  \end{cases}
\end{equation}

which is employed to compute the $T_i$ for each pixel $i$ of the study area as in:

\begin{equation}
  \label{eq:tair-nomix}
  T_i^{no \, mix} = T_{ref} + (1 - HM_i) \cdot UHI_{max}
\end{equation}

where $T_{ref}$ and $UHI_{max}$ are the rural reference temperature and the magnitude of the UHI effect respectively, as described above.
Finally, the $T_{air}$ values of each pixel $T_i^{no \, mix}$ are spatially averaged using a Gaussian function with a kernel radius $r$ defined by the user.

Following the calibration of the model to the same study area by Bosch et al. \cite{bosch2020spatially}, the values of the cooling capacity weights have been set to $w_{S} = 0.59$, $w_{AL} = 0.24$ and $w_{ET} = 0.17$, the distance over which green spaces have a cooling effect $d_{cool}$ and the air mixing radius $r$ have been set to 89.21 and 236.02~m respectively.
The $T_{ref}$ and $UHI_{max}$ values have been set based on the air temperature of 11 monitoring stations in the study area (see \autoref{fig:monitoring-stations}). More precisely, $T_{ref}$ has been set as the 9 p.m. air temperature measurement --- the moment of maximal UHI intensity in Switzerland \cite{burgst2019representing} --- of the station showing the lowest temperature value, and $UHI_{max}$ has been set as the difference between the 9 p.m. temperature measurement of the station showing the highest temperature value and $T_{ref}$.
% Given that the UHI effect in Switzerland reaches its maximal intensity around 9 p.m. \cite{burgst2019representing}, this study evaluates it based on the air temperature observations of that hour.
With the above definitions, a reference day for the simulations has been selected from the 2018-2019 period as the day showing the maximum $UHI_{max}$ with $T_{ref} > 20$. Such a date corresponds to July 27$^{th}$ 2018, with $T_{ref} = 20.60 \degree C$ and $UHI_{max} = 7.38 \degree C$.

\subsection*{Refining LULC classes based on tree cover and building density}
\label{sec:refin-lulc-class}

% To better account for the spatial heterogeneity of cities, a procedure to redefine the LULC classes from the cadastral survey has been designed, which further distinguishes the LULC classes depending on their proportional cover of both trees and buildings. Such reclassification is based on combining the 10~m raster LULC map with two 1~m binary raster masks, one for the tree canopy raster and another for the buildings. % , and consists of three main steps.
A procedure to redefine the LULC classes from the cadastral survey has been designed to distinguish the LULC classes depending on their proportional cover of both trees and buildings. The reclassification is achieved by combining the 10~m raster LULC map with two 1~m binary raster masks, one for the tree canopy raster and another for the buildings. % , and consists of three main steps.
The 1~m binary tree canopy mask has been derived from the SWISSIMAGE orthomosaic \cite{swisstopo2019swissimage}, by means of the Python library DetecTree \cite{bosch2020detectree}, which implements the methods proposed by Yang et al. \cite{yang2009tree}. On the other hand, the 1~m binary building mask has been obtained by rasterizing the buildings of the vector cadastral survey \cite{asitvd2018structure}.

% The reclassification procedure consists of three steps.
% % First, for each LULC class, the set of 10m pixels of such class is coupled with the tree canopy and building masks
% Firstly, each 10~m pixel is coupled with the tree canopy and building masks in order to respectively compute its proportion of tree and building cover.
% % Secondly, each LULC class is further refined into a set of subclasses depending on the joint distribution of tree cover and building cover of its 10~m pixels, e.g., the ``sidewalk'' LULC code might be further refined into ``sidewalk with low tree/low building cover'', ``sidewalk with low tree/high building cover'', ``sidewalk with high tree/low building cover'' and ``sidewalk with high tree/high building cover''. %  its set of 10~m pixels are % the 10m pixels of each LULC class are grouped into equally-spaced bins according to their proportion of tree cover
% Secondly, the set of 10~m pixels of each LULC class are grouped into a set of bins to form two histograms, one based on their proportion of tree cover and the other analogously for the building cover.
% % which groups the 10~m pixels of each LULC class into an increasing number of equally-spaced bins according to their proportion of tree cover, until adding an extra bin results in having a bin whose number of samples is lower than certain threshold.
% The number of bins is determined by means of an iterative procedure which distributes the 10~m pixels into an increasing number of equally-spaced bins until adding an extra bin results in having a bin whose number of samples is lower than certain threshold.
% After empirical exploration, a threshold value of 3~\% has been manually chosen with the aim of obtaining a set of refined LULC codes that can be interpreted easily.
% % cartesian product
% Finally, the two histograms are joined so that each LULC class is further refined into a set of subclasses, e.g., the ``sidewalk'' LULC code might be further refined into ``sidewalk with low tree/low building cover'', ``sidewalk with low tree/high building cover'', ``sidewalk with high tree/low building cover'' and ``sidewalk with high tree/high building cover''.
The reclassification procedure consists of three steps.
Firstly, each 10~m pixel is coupled with the tree canopy and building masks in order to respectively compute its proportion of tree and building cover.
Secondly, the set of 10~m pixels of each LULC class are grouped into a user-defined set of bins to form two histograms, one based on their proportion of tree cover and the other analogously for the building cover.
% cartesian product
Finally, the two histograms are joined so that each LULC class is further refined into a set of classes. For example, if two bins were used for both the tree and building cover, the ``sidewalk'' LULC code might be further refined into ``sidewalk with low tree/low building cover'', ``sidewalk with low tree/high building cover'', ``sidewalk with high tree/low building cover'' and ``sidewalk with high tree/high building cover''.
In the present work, four equally spaced bins (i.e., distinguishing 0-25\%, 25-50\%, 50-75\% and 75-100\% intervals) have been used to reclassify each LULC class according to both the tree and building cover.

In order to adapt the biophysical table of the InVEST urban cooling model to the reclassified LULC classes, the shade coefficients are computed as the midpoint of the bin interval of each level of tree cover (i.e., 0.125, 0.375, 0.625 and 0.875), whereas the albedo coefficients have been linearly interpolated based on the level of building cover (see \autoref{tab:biophysical-table}).

% \subsubsection{Calibration and evaluation of the model}

% The urban cooling model is calibrated to the study area following the approach proposed by Bosch et al. \cite{bosch2020spatially}, which consists in employing a simulated annealing procedure to find the combination of cooling capacity weights $w_{S}$, $w_{AL}$ and $w_{ET}$, the distance over which green spaces have a cooling effect $d_{cool}$ and the air mixing radius $r$ that best-fit the air temperature measurements from monitoring stations.

% In the present work, the calibration is based on the measurements of the same monitoring stations and the same dates as Bosch et al. \cite{bosch2020spatially}, yet . The coefficient of determination of the calibrated model is $R^2 = 0.90$, with a mean absolute error of $MAE = 1.19 \degree C$ and root-mean-square error of $RMSE = 1.45 \degree C$.

% \subsection{Relationship between the spatial pattern of tree canopy on urban heat mitigation}

% In order to explore the relationships between the

% The OLS, diagnostics for spatial dependece and spatial regression are performed with the PySAL library (v2.3.0) \cite{rey2010pysal}, whereas the Pearson and partial correlation analysis are performed respectively with the SciPy (v1.5.2) \cite{virtanen2020scipy} and Pingouin (v0.3.7) \cite{vallat2018pingouin} libraries.

\subsection*{Generation of urban greening scenarios}

% One of the main advantages of using models based on biophysical mechanisms is that they allow for reasonable
% Departing from the refined LULC map, a set of synthetic scenarios will be simulated by altering the LULC classes of certain pixels in a way that corresponds to reasonable transformations that could occcur in urban areas.
% More precisely, in order to mimic scenarios of urban greening, pixels whose base LULC class corresponds to roads, sidewalks, traffic islands and other impervious surfaces will be changed to the LULC code that has the same base class but with the highest tree cover, e.g.,  pixels of a post-refinement class ``sidewalk with low tree/low building cover'' would be changed to ``sidewalk with high tree/low building cover''.
Departing from the refined LULC map, a set of urban greening scenarios will be simulated by altering the LULC classes of certain pixels in a way that corresponds to reasonable transformations that could occcur in urban areas.
More precisely, pixels whose base LULC class corresponds to ``building'', ``road, path'', ``sidewalk'', ``traffic island'', ``other impervious'' and ``garden'' will be changed to the LULC code that has the same base class but with the highest tree cover, e.g.,  pixels of a post-refinement class ``sidewalk with low tree/low building cover'' would be changed to ``sidewalk with high tree/low building cover''.
% limit such that the proportion of tree canopy cover and the proportion of building cover (i.e., "shade" + "bulding_intensity") does not amount to more than 100~\%
In order to ensure that such an increase of the tree canopy cover is performed only where the existing urban fabric permits it, the transformed pixels . First, the transformations might only take place in pixels where the proportion of building cover is under 25\%, i.e., there is a 75\% of the pixel area which could be occupied by a tree crown. Secondly, pixels of the ``road, path'' class might only be transformed when they are next to a pixel of a different class, which prevents increasing the tree canopy cover in pixels that are in the middle of a road (e.g., a highway).

% with two main objectives. The first objective is to use the scenarios to explore the relationships between spatial patterns in both composition and configuration of trees and the distribution of air temperature simulated with the urban cooling model. The second objective is to inform policy-making.
% In order to explore the relationships between spatial pattern of the tree canopy in each scenario and its respective distribution of air temperature simulated with the urban cooling model.
% In order to quantify spatial pattern of the tree canopy of each scenario, a set of spatial metrics from landscape ecology \cite{o1988indices,mcgarigal2012fragstats} are computed for the pixels whose post-refinement LULC class has a high tree canopy cover.  % over a certain threshold.
For each scenario, the spatial pattern of the tree canopy is quantified by means of a set of spatial metrics from landscape ecology \cite{o1988indices,mcgarigal2012fragstats}, which are computed for the pixels whose post-refinement LULC class has a tree canopy cover over 75\% \footnote{Following the advice given by the directorate of resoures and natural heritage in the Canton of Vaud (DGE-DIRNA), the threshold over which a pixel is considered to have a high tree canopy cover has been set to 75~\%, which corresponds to placing trees of a spheric crown with a 5~m radius spaced 10~m from one another so that they form a continuous canopy.}.
Based on other studies that explore the relationship between the spatial of tree canopy and UHIs, four spatial metrics have been chosen to quantify both the composition and oconfiguration of the tree canopy, which are listed in \autoref{tab:selected-metrics}. The proportion of landscape (PLAND) of pixels with high tree canopy cover serves to quantify the composition aspects, while the configuration is quantified by means of the mean patch size (MPS), edge density (ED) and the mean shape index (MSI) of patches of high tree canopy cover. The four metrics have been computed with the Python library PyLandStats \cite{bosch2019pylandstats}.

\begin{table}[!h]
  \begin{adjustwidth}{-.4\textwidth}{0cm} % Comment out/remove adjustwidth environment if table fits in text column.
    % \footnotesize % Font size can be changed to match table content. Recommend 10 pt.
    \caption{\label{tab:selected-metrics}Selected landscape metrics. A more thorough description can be found in the documentation of the software FRAGSTATS v4 \cite{mcgarigal2012fragstats}}
    \renewcommand{\arraystretch}{1.5} % vertical spacing
    \begin{center}
      \begin{tabular}{p{.15\textwidth} p{.4\textwidth} p{.76\textwidth}} 
        \toprule
        \textbf{Category} & \textbf{Metric name} & \textbf{Description} \\
        \midrule
        Composition & Percentage of landscape (PLAND) & Percentage of landscape, in terms of area, occupied by pixels with high tree canopy cover \\
        Configuration & Mean patch area (AREA\_MN) & Average size (in hectares) of the patches formed by pixels with high tree canopy cover \\
        & Mean shape index (SHAPE\_MN) & Average shape index of the patches formed by pixels with high tree canopy cover \\
        & Edge density (ED) & Sum of the lengths of all edge segments between pixels with high tree canopy cover an other pixels, per area unit (in m/hectare) \\    
        \bottomrule  
      \end{tabular}
    \end{center}
  \end{adjustwidth}
\end{table}

\subsubsection*{Simulation of urban greening by randomly increasing the proportion of tree cover}
With the aim of evaluating the effect of the proportion of tree cover on air temperature, a set of scenarios is generated by transforming a 25, 50, 75 and 100\% of the pixels where the tree canopy cover can be increased.
For each of the aforementioned percentages, 10 scenarios will be generated by randomly sampling from the candidate pixels until the desired proportion of changed pixels is matched. Therefore, the temperature maps for each of the percentage of changed pixels will be computed by averaging the 10 respective simulated runs.

\subsubsection*{Simulation of distinct urban greening approaches by clustering and scattering trees}

% To expore the relationship between
In order to obtain a greater variety of spatial patterns of tree canopy cover and further explore the relationship between the spatial configuration of tree canopy and UHIs, two additional sets of scenarios will be generated.
% Like the first set of scenarios described above, these two additional sets are also produced by transforming a 25, 50, 75 and 100\% of the pixels where the tree canopy can be increased.
% Nevertheless, in these two sets the transformed pixels are not randomly sampled but selected according to the number of pixels with high tree canopy cover (i.e., > 75\%) found in their Moore neighborhood. Therefore,
In these sets of scenarios, the transformed pixels are not randomly sampled but selected according to the number of pixels with high tree canopy cover (i.e., greater than 75\%) found in their Moore neighborhood.
Therefore, the first set of scenarios pixels with higher number of high tree canopy cover neighbors are transformed first, which intends to spatially cluster pixels of high tree canopy cover. Conversely, the second set of scenarios intends to spatially scatter pixels of high tree canopy cover by giving precedence to pixels with lower number of high tree canopy neighbors.
To explore the difference between the two approaches with more detail, these two sets of scenarios will be produced by transforming a 12.5, 25, 37.5, 50, 62.5, 75, 87.5 and 100\% of the pixels where the tree canopy can be increased.
Like the previous set of scenarios, the temperature maps will be computed by averaging the results of 10 simulation funs for each configuration, i.e., combination of proportion of changed pixels and spatial approach of either clustering or scattering the transformed pixels.

\section*{Results}

% \subsection{Performance of the urban cooling model}

% \subsection{Relationship between the spatial pattern of tree canopy on urban heat mitigation}

% The coefficients estimated with the OLS regression are listed in \autoref{tab:selected-metrics}. Although the values of the Moran I \cite{moran1950notes} suggest a strong spatial autocorrelation, the use of a spatial regression model does not improve the coefficient of determination nor change the sign of any coefficient at any scale (see \ref{sec:spatial-relationships}). Therefore, the remainder of the article will only consider the OLS regression.
% In line with the results of Zhou et al. \cite{zhou2017effects}, the coeffient of determination increases with the scale, whereas the Moran I and the Akaike information criterion (AIC) \cite{akaike1973information} decrease, suggesting that the linear regression model performs better at larger scales.

% \begin{table}
%   \caption{\label{tab:ols-regression} Results of the OLS regression.}
%   \scriptsize % Font size can be changed to match table content. Recommend 10 pt.
%   \begin{center}
%     \begin{tabular}{ p{.065\textwidth} p{.11\textwidth} p{.11\textwidth} p{.11\textwidth} p{.11\textwidth} p{.11\textwidth} p{.055\textwidth} p{.07\textwidth} p{.07\textwidth} }
%       \toprule
%       Scale & PLAND & AREA\_MN & SHAPE\_MN & ED & DIST & R$^2$ & Moran & AIC\\
%       \midrule
%       120~m & -0.013741** & 0.375819** & -0.169583** & 0.000688** & -0.000119** & 0.414 & 0.872** & 9472.641 \\
%       & \textbf{-0.353207} & \textbf{0.130888} & \textbf{-0.055450} & \textbf{0.102536} & \textbf{-0.347371} \\
%       360~m & -0.016759** & 0.111463** & -0.365982** & 0.001481** & -0.000113** & 0.459 & 0.796** & 1335.884 \\
%       & \textbf{-0.340134} & \textbf{0.166164} & \textbf{-0.145221} & \textbf{0.175597} & \textbf{-0.337706} \\      
%       600~m & -0.023965** & 0.265071** & -0.635437** & 0.002251** & -0.000105** & 0.491 & 0.678** & 501.365 \\
%       & \textbf{-0.436498} & \textbf{0.248360} & \textbf{-0.202908} & \textbf{0.240449} & \textbf{-0.313095} \\
%       840~m & -0.029608** & 0.295937* & -0.529756** & 0.003047** & -0.000098** & 0.533 & 0.572** & 255.563 \\
%       & \textbf{-0.489575} & \textbf{0.180218} & \textbf{-0.161129} & \textbf{0.291127} & \textbf{-0.296891} \\
%       1080~m & -0.023499** & 0.233338 & -1.413028** & 0.002752** & -0.000106** & 0.564 & 0.486** & 154.011 \\
%       & \textbf{-0.360574} & \textbf{0.139134} & \textbf{-0.190122} & \textbf{0.233989} & \textbf{-0.311524} \\      
%       \bottomrule
%     \end{tabular}
%   \end{center}
%   ** $P < 0.01$, * $P < 0.05$ (2-tailed)
% \end{table}

% The standardized coefficients suggest that proportion of tree canopy cover is the feature that has most heat mitigation effect at all scales, followed by the distance to the city center. Among the configuration metrics, the edge density has the most significant effect on air temperature, which suggest that a more edgy tree canopy pattern contributes to higher temperatures. A similar positive effect is observed for the mean patch size, nonetheless it becomes less significant at larger scales. Finally, the mean shape index has a negative effect on air temperature which indicates that more irregular tree canopy patches contribute to greater heat mitigation.

% % \begin{table}
% %   \footnotesize % Font size can be changed to match table content. Recommend 10 pt.
% %   \begin{center}
% %     \begin{tabular}{ p{.07\textwidth} p{.11\textwidth} p{.11\textwidth} p{.11\textwidth} p{.11\textwidth} p{.11\textwidth} p{.07\textwidth} p{.08\textwidth} }
% %       \toprule
% %       Scale & PLAND & AREA\_MN & SHAPE\_MN & ED & DIST & R$^2$ & AIC\\
% %       \midrule
% %       120~m & -0.027518 & 1.075121 & -0.340077 & 0.001006 & -0.000118 & 0.381 & inf \\
% %       360~m & -0.020581 & 0.088502 & -0.206741 & 0.001568 & -0.000112 & 0.449 & 1664.495 \\
% %       600~m & -0.003655 & 0.033896 & -0.150886 & 0.000321 & -0.000131 & 0.357 & -248.053 \\
% %       840~m & -0.007555 & 0.073376 & -0.163887 & 0.000450 & -0.000123 & 0.411 & -39.204 \\
% %       1080~m & -0.014000 & 0.146802 & -0.678353 & 0.000938 & -0.000103 & 0.535 & 114.427 \\
% %       \bottomrule
% %     \end{tabular}
% %     \caption{\label{tab:spatial-regression} Results of the spatial regression.}
% %   \end{center}
% % \end{table}

% The correlation coefficients for the Pearson correlation and the partial correlation analysis are listed in \autoref{tab:correlation-coefficients}. The distance to the city center is the feature that is most negatively correlated to air temperature, and its effect increases when controlling for the proportion of tree cover. Each of the configuration metrics shows a distinctive response to the partial correlation analysis. The mean patch size becomes shows a clear decrease on its negative influence on air temperature when controlling for the proportion of tree cover, even losing the $P < 0.05$ significance level at the largest scale. In sharp contrast, the mean shape index increases its negative effect on air temperature, whereas the edge density changes from a positive correlation with air temperature to a negative effect with a similar level of significance. Contrarily to the OLS regression above, the results of the partial correlation suggest that increasing the edges between tree and non-tree pixels contributes to urban heat mitigation. On the other hand, when controlling for the effects of the configuration metrics, the proportion of tree cover shows an increase on its negative effect on air temperature. Finally, with very few exceptions, the correlation coefficients increase with the scale.

% \begin{table}
%   \caption{\label{tab:correlation-coefficients} Correlation coefficients. The bold rows show the coefficients of the partial correlation analysis.}
%   \footnotesize % Font size can be changed to match table content. Recommend 10 pt.
%   \begin{center}
%     \begin{tabular}{ p{.075\textwidth} p{.13\textwidth} p{.13\textwidth} p{.13\textwidth} p{.13\textwidth} p{.13\textwidth}}
%       \toprule
%       Scale & PLAND & AREA\_MN & SHAPE\_MN & ED & DIST \\
%       \midrule
%       120~m & -0.335816** & -0.326044** & -0.203751** & -0.021708 & -0.542007** \\
%       & \textbf{-0.201833}** & \textbf{-0.053347}** & \textbf{-0.140807}** & \textbf{-0.024849}* & \textbf{-0.575068}** \\
%       360~m & -0.304223** & -0.230111** & -0.293065** & 0.090014* & -0.545207** \\
%       & \textbf{-0.353835}** & \textbf{-0.038458} & \textbf{-0.299635}** & \textbf{-0.145022}** & \textbf{-0.595347}** \\      
%       600~m & -0.340180** & -0.334422** & -0.343264** & 0.111613* & -0.535269**  \\
%       & \textbf{-0.398365}** & \textbf{-0.129310}* & \textbf{-0.374636}** & \textbf{-0.177379}** & \textbf{-0.593629}** \\
%       840~m & -0.368275** & -0.392994* & -0.343015** & 0.114636 & -0.544322** \\
%       & \textbf{-0.428976}** & \textbf{-0.180682}* & \textbf{-0.457676}** & \textbf{-0.173731}* & \textbf{-0.606605}** \\
%       1080~m & -0.342848** & -0.346385 & -0.354632** & 0.236678** & -0.569577**  \\
%       & \textbf{-0.428303}** & \textbf{-0.177690} & \textbf{-0.484208}** & \textbf{-0.215491}* & \textbf{-0.650891}** \\      
%       \bottomrule
%     \end{tabular}
%   \end{center}
%   ** $P < 0.01$, * $P < 0.05$ (2-tailed)
% \end{table}

% \subsection{Evaluation of the greening scenarios}

\subsection*{LULC change}

The relationship between the number of pixels by LULC class changed and the proportion of pixels changed to its corresponding LULC class with high tree canopy cover is shown in \autoref{fig:greening-lulc-change}.
\begin{figure}
  \centering
  \includegraphics[width=.8\textwidth]{figures/greening-lulc-change} 
  \caption{\label{fig:greening-lulc-change} Number of pixels by its original LULC class changed to its corresponding LULC class with high tree canopy cover for a proportion of pixels changed of 25, 50, 75 and 100\%. The lines at the top of the bars represent the 95\% confidence intervals.}
\end{figure}
Changing a 25, 50, 75 and 100\% of the potential pixels corresponds to a total number of pixels changed of 118880, 237760, 356640 and 475520 respectively.
When increasing the tree canopy in all the possible pixels, a 61.4\% of the pixels correspond to the ``garden'' LULC class, followed by ``road, path'', ``building'', ``other impervious'', which account for a a 18.01, 10.81 and 7.69\% respectively. Finally, the LULC classes of ``sidewalk'' and ``traffic island'' constitute only a 1.67 and 0.3\% of the  pixels where the tree canopy can be increased.
% CI are of course, inexistent when changing all the possible pixels
% The confidence intervals are very small for all the proportions of pixels changed, which reflects the constraints imposed by the existing urban fabric.

\subsection*{Effect of the proportion of tree cover}

% A set scenario LULC maps generated with increasing proportion of pixels changed to its corresponding LULC code with high tree canopy cover, its corresponding simulated air temperatuers and heat mitigation are shown in \autoref{fig:scenarios-prop} (see \ref{sec:scenarios-prop}).
The maps for four scenarios generated by changing an increasing proportion of pixels to its corresponding LULC code with high tree canopy cover are shown in \autoref{fig:scenarios-prop}.
% \begin{figure}
%   \begin{subfigure}{\textwidth}
%     \centering
%     \includegraphics[width=\textwidth]{figures/scenarios-prop-lulc.png}
%   \end{subfigure}
%   \begin{subfigure}{\textwidth}
%     \centering    
%     \includegraphics[width=\textwidth]{figures/scenarios-prop-T.png}  
%   \end{subfigure}
%   \begin{subfigure}{\textwidth}
%     \centering    
%     \includegraphics[width=\textwidth]{figures/scenarios-prop-mitigation.png}
%   \end{subfigure}
%   \caption{\label{fig:scenarios-prop} Scenario raster maps with an increasing (from left to right) proportion of the pixels changed to its corresponding LULC code with high tree canopy cover. The rows respectively feature the pixels where the LULC is changed (top), the simulated air temperature (middle) and the simulated heat mitigation (bottom).}
% \end{figure}
\begin{figure}
  \begin{adjustwidth}{-.4\textwidth}{0cm}
    \includegraphics[width=\linewidth]{figures/scenarios-prop-lulc.png}
    \includegraphics[width=\linewidth]{figures/scenarios-prop-T.png}
    \includegraphics[width=\linewidth]{figures/scenarios-prop-mitigation.png}
    % \begin{subfigure}{\linewidth}
    %   \includegraphics[width=\textwidth]{figures/scenarios-prop-lulc.png}
    % \end{subfigure}
    % \begin{subfigure}{\linewidth}
    %   \includegraphics[width=\textwidth]{figures/scenarios-prop-T.png}  
    % \end{subfigure}
    % \begin{subfigure}{\linewidth}
    %   \includegraphics[width=\textwidth]{figures/scenarios-prop-mitigation.png}
    % \end{subfigure}
    % \captionsetup{width=\linewidth, margin={0cm,-.4\textwidth}}
    \caption{\label{fig:scenarios-prop} Scenario raster maps with an increasing (from left to right) proportion of the pixels changed to its corresponding LULC code with high tree canopy cover. The rows respectively feature the pixels where the LULC is changed (top), the simulated air temperature (middle) and the simulated heat mitigation (bottom).}  
  \end{adjustwidth}
\end{figure}
When changing a 25, 50, 75 and 100\% of the candidate pixels, the maximum temperature $T$ for the reference date, i.e., 26.05$\degree C$, is progressively reduced to 25.66, 25.27, 24.87 and 24.47 $\degree C$ respectively, while the magnitude of maximum heat mitigation $T - T_{obs}$ increases from 0.51, 1.047, 1.54 and 2.05$\degree C$ respectively.
The largest heat mitigation magnitudes occur in the most urbanized parts, which are located along the main transportation axes.
The relationship between the proportion of candidate pixels transformed and the simulated distribution of air temperature can be approximated by a straight line with a negative slope (see \autoref{fig:scenarios-prop-regplot} and \autoref{fig:scenarios-prop-hists} for more details of this relationship).
% The strength of such a fit suggests that the constraints on the urban heat mitigation imposed by the existing urban fabric .

\subsection*{Effect of the spatial configuration of tree cover}

The relationships between the landscape metrics computed for each scenario sample and the corresponding average simulated temperature are displayed in \autoref{fig:scenarios-config}.
\begin{figure}
  \begin{adjustwidth}{-.4\textwidth}{0cm}  
    \includegraphics[width=.99\linewidth]{figures/scenarios-config}
    % \captionsetup{width=\linewidth, margin={0cm,-.4\textwidth}}
    \caption{\label{fig:scenarios-config} Relationship between landscape metrics and the average simulated temperature $\overline{T}$ for each scenario sample, colored to distinguish scenarios simulated with the ``cluster'' and ``scatter'' approaches.}    
  \end{adjustwidth}
\end{figure}
% The values of PLAND, which are directly related to the proportion of tranformed candidate pixesl
The proportion of landscape (PLAND) occupied by pixels with high tree canopy cover range from 17.26 to 53.37\%. As a composition metric, PLAND is directly related to the proportion of transformed candidate pixels, and the start and end of the PLAND value range correspond to transforming 0 and 100\% of the candidate pixels respectively.
The relationship between PLAND and the average simulated temperature of each scenario $\overline{T}$ is almost perfectly linear, however, for the same PLAND values, scattering the pixels with high tree canopy cover instead of clustering them consistently leads to lower $\overline{T}$.
% With the exception of the mean shape index,

Regarding the configuration metrics, the values of the mean patch area (AREA\_MN) show that clustering the transformed pixels leads to larger patches of high tree canopy cover than the scattering approach. Conversely, scattering the transformed pixels of high tree canopy cover results in higher edge density (ED) than clustering them. 
The average simulated temperature $\overline{T}$ is overall negatively correlated with AREA\_MN, which suggests that for the same amount of high tree canopy pixels, large patches provide lower heat mitigation.
On the other hand, configurations with the same proportion of high tree canopy pixels show lower $\overline{T}$ for larger values of ED, which suggests that edge effects between artificial patches and patches of high tree canopy contribute to greater heat mitigation.
% TODO: Nevertheless, the relationship between $\overline{T}$ shows two distinct regimes. For large values of PLAND,
% TODO: Such a relationship is likely due to spatial constraints imposed by the existing urban fabric .
% TODO: Samples with the same proportion of changed pixels show practically the exact same values for the mean patch area and edge density, whereas a slight differences can be appreciated in the mean shape index values.
Finally, the mean shape index (SHAPE\_MN) shows no discernable relationship with $\overline{T}$.

\subsection*{Effects on human exposure}

% The fraction of the population exposed to temperatures higher than certain thresholds for a set of scenarios with increasing proportion of tree cover
The relationship between human exposure to air temperatures higher than 21, 22, 23, 24, 25 and 26$\degree$C and the proportion of pixels transformed to their respective high tree canopy cover class is shown in \autoref{fig:human-exposure}.
\begin{figure}[ht]
  \centering
  \includegraphics[width=.98\textwidth]{figures/human-exposure}
  \caption{\label{fig:human-exposure} Population exposed to temperatures higher than 21, 22, 23, 24, 25 and 26$\degree$C respectively, when transforming 0, 25, 50, 75 and 100\% of candidate pixels to its corresponding LULC code with high tree canopy cover. The left plot shows the number of dwellers. The right plot shows the difference between the number of dwellers exposed to each temperature in the ``cluster'' and ``scatter'' approaches. The lines at the top of the bars represent the 95\% confidence intervals.}
\end{figure}
% The number of dwellers exposed to temperatures higher than 22$\degree$C diminishes as the number of pixels transformed to high tree canopy classes increases. 
The number of dwellers exposed to temperatures higher than 21$\degree$C does not practically change (it is reduced from 274754 to 274748 when the proportion of transformed pixels is of 75\%), whereas for temperatures higher than 22$\degree$C, it diminishes from 269254 to 268601, 267683, 266518 and 264125 when the proportion of transformed pixels is of 25, 50, 75 and 100\% respectively, representing a relative share of 97.25, 97.02, 96.69 96.27 and 95.41\% of the population of the study area.
Such a decline progressively becomes more notable as temperatures increase.
For temperatures over 23$\degree$C, the corresponding decreases start from 254653 to 250392, 243063, 231065 and 210802, representing a 91.98 to 90.44, 87.80, 83.46 and 76.14\% of the population, while for temperatures over 24$\degree$C 217051, 200406, 164908, 103889 and 31905, respectively representing a 78.40, 72.39, 59.57, 37.53 and 11.52$\degree$C.
Finally, the number of dwellers exposed to temperatures over 25$\degree$C decreases from 132641 to 69162 and 15866 for proportions of transformed pixels of 25 and 50\% respectively, representing a 47.91, 24.98 and 5.73\% of the total population, and becomes 0 after that, whereas the totality 2508 dwellers originally exposed to temperatures over 26$\degree$C, which represents a 0.91\% of the population, do no longer meet such temperatures after transforming a 25\% of the candidate pixels.

% The differences between the spatially scattering or scattering the transformed pixels are most notable when considering the number of dwellers exposed to temperatures higher than 24$\degree$C and 25$\degree$C.
Regarding the spatial configurations, scattering the transformed pixels instead of clustering them with other high tree canopy pixels can result in a reduction of the 3238, 9597 and 15323 dwellers exposed to temperatures higher than 24$\degree$C for the same proportions of transformed pixels of 25, 50 and 75\%, which respectively represents a 1.17, 3.47 and 5.53\% of the total number of dwellers. For temperatures higher than 25$\degree$C, the analogous reduction is of 20708 and 22748 dwellers when transforming a 25 and 50\% of the pixels, which represents a 7.48 and 8.22\% of the population.

\section*{Discussion}

The scenarios simulated in this study represent a new form of exploring the heat mitigation potential provided by the urban tree canopy cover.
The results map locations where the tree canopy cover in the urban agglomeration of Lausanne can be increased, and shows that such changes can result in urban nighttime temperatures that are up to 2$\degree$C lower.
Additionally, the number of dwellers exposed to nighttime temperatures over 24$\degree$C can be reduced from 217051 to 31905, which respectively represents a 78.40 and 11.52\% of the total population in the study area.

% 1. EFFECTS OF THE SPATIAL CONFIGURATION
% % The proposed simulation approach presents several major advantages when compared to statistical analyses.
% Overall, the simulated scenarios in Lausanne indicate that given the same proportion of tree canopy cover, a scattered configuration might result in more effective urban heat mitigation than a clustered one, both in terms of the average temperature and the number of urban dwellers exposed to the highest tempetratures.
% % might have the spatial configuration of the tree canopy might indeed have a significant effect on the spatial pattern of air temperature, yet it seems that such an effect is very small when compared to the effect of the proportion of tree cover
% Nevertheless, the results suggest that effect of the spatial configuration is secondary when compared to the effect of the composition, i.e., proportion of tree cover.
% % Trees contribute to urban heat mitigation in two primary ways, namely the shading and evapotranspiration processes. Therefore, the effect of the spatial configuraiton of trees on its cooling effectiveness
% The effect of the spatial configuration of trees on its urban cooling effectiveness depends on how it affects the shading and evapotranspiration processes. Such a relationship is know to be highly complex and dependent on the ecological context \cite{zhou2017effects}, tree species \cite{jiao2017patch}
The simulated scenarios in Lausanne indicate that given the same proportion of tree canopy cover, a scattered configuration might result in more effective urban heat mitigation than a clustered one, which is in line with previous studies in humid climates \cite{kong2014effects,estoque2017effects,zhou2017effects,yu2018strong,nastran2019urban}.
Nevertheless, the results suggest that effect of the spatial configuration is secondary when compared to the effect of the composition, i.e., proportion of tree cover.
% Overall, the effect of the spatial configuration of trees on its urban cooling effectiveness depends on how it affects the shading and evapotranspiration processes, which is known to be highly complex relationship that is strongly mediated by the tree species as well as the local climatic and environmental conditions \cite{zhou2017effects,jiao2017patch,yu2018strong,wang2020significant}. Furthermore, the cooling effects of tree canopy cover are scale-dependent \cite{li2013relationship,estoque2017effects,yan2019testing,wang2020significant,terfa2020spatial}.
Overall, the effect of the spatial configuration of trees on its urban cooling effectiveness depends on how it affects the shading and evapotranspiration processes.
Such a relationship is known to be strongly mediated by the tree species, background climatic and environmental conditions as well as the spatial scale \cite{li2013relationship,estoque2017effects,zhou2017effects,jiao2017patch,yu2018strong,yan2019testing,wang2020significant,terfa2020spatial}.
In order to better represent the effects of the spatial configuration of the tree canopy, the InVEST urban cooling model must be further validated with experiments at the neighborhood scale to ensure that it provides an appropriate city-scale depiction of how the urban heat mitigation mechanisms operate at finer scales.

% 2. NON-LINEARITIES AND THRESHOLDS
% The proposed approach provides an effective interface to spatially evaluate mitigation strategies. Further refinement of the urban cooling model (i.e., including non-linearities) might enable further explorations (e.g., thresholds).
% Multi-scale perspective \cite{wang2020spatial}
% gains are orthogonal to those of green roofs
% We also need to compare the urban cooling model to more fine-grained models (e.g., three-dimensional models at the neighborhood scale) to find out whether the urban cooling model provides an adequate averaging of the neighborhood scale behaviour. We have the required high-resolution data for most parts (e.g., cadastre, tree canopy) but not high-resolution air temperature measurements (in space and time). Note that LST is not good enough
% While the InVEST urban cooling model provides an effective method to simulate urban heat mitigation from LULC rasters at the urban agglomeration scale \cite{bosch2020spatially}, it must be further validated with experiments at the neighborhood scale to ensure that it provides an appropriate upscale representation of how the urban heat mitigation mechanisms operate at finer scales.
On the other hand, the InVEST urban cooling models presents limitations regarding the simplified and homogeneous way in which the air is mixed, as well as the cooling effects of large green spaces \cite{sharp2020invest,bosch2020spatially}.
As a result, the relationship between the proportion of tree canopy cover and the magnitude of the urban heat mitigation reported in this work is practically linear, and the temperature differences between spatially clustering or scattering the new tree canopy cover are limited.
Nonetheless, in complex terrains such as the Lausanne agglomeration, models with uniform weighting of space show considerable deviations from the observed spatial patterns of air temperature \cite{frei2014interpolation}. Moreover, the cooling effects of large green spaces have been found to be non-proportional to their area and shape complexity \cite{chen2014effect,bao2016assessing,du2017quantifying}.
% Including non-linearities of this kind may not only make the urban cooling model more realistic but also more useful for urban planning, e.g., by suggesting thresholds in the relatinship between the proportion of tree canopy cover and the simulated heat mitigation that ensure ``optimal'' spending.
Improving how these non-linear components are represented in the InVEST urban cooling model could enhance not only its validity, but also its value to urban planning by identifying thresholds and regime changes in the cooling efficiency of additional tree planting.
% MAYBE: while there is scope for improvement, the current implementation of the model suggests that there are thresholds regarding human exposure.
% MAYBE: virtual explorations: ``A concept that emerges quite naturally from such purposive bottom-up actions is the idea that to explore good planning and design in cities, computer models should be set up in a laboratory-like context in which the focus is on exploration of different patterns which attempt to reach different goals, laboratories in which models are available in wide area mode, across the web in a form that many people can experiment with.'' \cite[page 115]{batty2010complexity}, \cite{de2010planner,bosch2019addressing}

% 3. EVALUATE FUTURE SCENARIOS, COMPARE SPRAWL VS INFILL
% Additionally, we can use the approach to compare distinct urban expansion scenarios (e.g., compact, sprawl) \cite{lemonsu2015vulnerability,yang2016contrasting}, also landscape homogenization along the urban-rural gradient \cite{wang2020spatial}.
Despite the limitations noted above, a major advantage of the proposed approach is that it can be used to evaluate urban heat mitigation of synthetic scenarios.
The simulations presented in this article focus on spatially exploring the effects of an increase of the tree canopy cover, yet there is room for much more experimentation of this kind.
% MAYBE: landscape ecology principals/network planning to ensure connectivity and the like? see \cite[page 766]{haaland2015challenges}
For instance, in line with recent studies \cite{lemonsu2015vulnerability,yang2016contrasting,trimmel2019thermal}, the approach could be applied to examine the impact of distinct urbanization scenarios such as densification and urban sprawl on air temperature and human thermal comfort, under current conditions as well as future climate estimates, e.g., by changing the $T_{ref}$ or $UHI_{max}$ parameters.
Similarly, InVEST urban cooling model might be coupled with models of LULC change such as cellular automata in order to assess not only which scenarios are most desirable in terms of urban heat mitigation, but also which planning strategies might lead to them \cite{silva2008strategies,bosch2019addressing}.
% Also to evaluate projected LULC maps (e.g., outputs of cellular-automata simulations \cite{bosch2019addressing})
The spatiotemporal patterns of LULC change observed during the last 40 years in the Lausanne agglomeration have been characterized by infilling development and a progressive coalescensce of the artificial surfaces in its inner ring \cite{bosch2020spatiotemporal}.
% Overall decrease of urban green space in European cities due to infill development \cite{haaland2015challenges}.
Therefore, such an infilling trend urges for careful evaluation of the beneficial ecosystem services provided by urban green spaces, which should be balanced against the adverse consequences of urban sprawl \cite{haaland2015challenges,artmann2019urban}.

% \section*{Conclusion}
% TODO: conclusions

\section*{Supporting Information}

%These commands reset the figure/table counter and add "S" to the figure/table caption (e.g. "Figure S1"/"Table S1").
\setcounter{figure}{0}
\renewcommand{\thefigure}{S\arabic{figure}}
\setcounter{table}{0}
\renewcommand{\thetable}{S\arabic{table}}

\section*{Data}

\subsection*{Biophysical table}
\label{sec:biophysical-table}

% Biophysical table (before the reclassification), as comma-separated values (CSV).
% \url{https://github.com/martibosch/lausanne-heat-islands/blob/master/data/raw/biophysical-table.csv}
The biophysical table for the LULC codes (before the reclassification) is shown in \autoref{tab:biophysical-table}. The crop and water coefficients are based on Allen et al. \cite{allen1998crop}, while rock, soil and urban coefficients are derived from the results of Grimmond and Oke \cite{grimmond1999evapotranspiration} in the city of Chicago. Given that the evapotranspiration of the vegetation and crops is subject to seasonal changes in temperate zones such as Switzerland \cite{allen1998crop}, the values that correspond to the mid-season estimation (June to August) in \cite{nistor2016mapping}.
The albedo values are based on the work of Steward et al. \cite{stewart2012local}.
The shade column, which represents the proportion of tree cover of each LULC class, is computed after the reclassification procedure described in section ``\nameref{sec:refin-lulc-class}''. % \autoref{sec:refin-lulc-class}.

\begin{table}[!h]
  \begin{adjustwidth}{-.1\textwidth}{0cm}
    \caption{\label{tab:biophysical-table} Biophysical table (before the reclassification). The source comma-separated value (CSV) file used in the computational workflow is available at \url{https://github.com/martibosch/lausanne-heat-islands/blob/master/data/raw/biophysical-table.csv}.}
    \begin{center}
      \begin{tabular}{ c p{.28\textwidth} p{.16\textwidth} c c c }
        \toprule
        LULC code & Description & Case & $K_c$ & Albedo & Green area \\
        \midrule
        0 & building & artificial & 0.4 & 0.1-0.25 & 0 \\ % 0.142
        1 & road, path & artificial & 0.35 & 0.15 & 0 \\ % 0.141
        2 & sidewalk & artificial & 0.35 & 0.15 & 0 \\
        3 & traffic island & artificial & 0.35 & 0.15 & 0 \\
        4 & rail & artificial & 0.35 & 0.15 & 0 \\
        5 & airfield & artificial & 0.4 & 0.2 & 0 \\
        6 & pond & water & 0.45 & 0.15 & 0 \\
        7 & other impervious & artificial & 0.36 & 0.15 & 0 \\
        8 & field, meadow, pasture & vegetation & 0.9 & 0.2 & 1 \\
        9 & vineyards & vegetation & 0.7 & 0.2 & 1 \\
        10 & other intensive farming & vegetation & 1.05 & 0.2 & 1 \\
        11 & garden & artificial & 0.32 & 0.2 & 1 \\
        12 & wetland & water & 0.45 & 0.1 & 1 \\
        13 & other green & vegetation & 0.45 & 0.2 & 1 \\
        14 & backwater & water & 0.65 & 0.05 & 1 \\
        15 & water course & water & 0.65 & 0.05 & 0\\
        16 & reed & water & 0.45 & 0.1 & 1\\
        17 & dense forest & vegetation & 1.5 & 0.15 & 1 \\
        18 & densely wooded pasture & vegetation & 1.15 & 0.15 & 1 \\
        19 & open wooded pasture & vegetation & 1.15 & 0.2 & 1 \\
        20 & other wooded & vegetation & 1.15 & 0.15 & 1\\
        21 & bare rocks & rock and soil & 0.2 & 0.25 & 0 \\
        22 & glacier & water & 0.52 & 0.1 & 0 \\
        23 & sand & rock and soil & 0.3 & 0.25 & 0 \\
        24 & gravel pit & artificial & 0.36 & 0.25 & 0 \\
        25 & other non-vegetated & artificial & 0.36 & 0.15 & 0 \\
        \bottomrule
      \end{tabular}
      % \caption{\label{tab:evapotranspiration-coefficients}
      % Evapotranspiration coefficients for each LULC class (see \nameref{table-biophysical}). Crop and water coefficients are based on Allen et al. \cite{allen1998crop}, while rock, soil and urban coefficients are derived from the results of Grimmond and Oke \cite{grimmond1999evapotranspiration} in the city of Chicago. Given that the evapotranspiration of the vegetation and crops is subject to seasonal changes in temperate zones such as Switzerland \cite{allen1998crop}, the values that correspond to the mid-season estimation (June to August) in \cite{nistor2016mapping}.
      % The albedo values are based on the work of Steward et al. \cite{stewart2012local}.
      % The source comma-separated value (CSV) file used in the computational workflow is available at \url{https://github.com/martibosch/lausanne-heat-islands/blob/master/data/raw/biophysical-table.csv}.
      % }
    \end{center}
    % \vspace{.5em}
    % \textbf{Table S3.} Biophysical table. The crop and water coefficients are based on Allen et al. \cite{allen1998crop}, while rock, soil and urban coefficients are derived from the results of Grimmond and Oke \cite{grimmond1999evapotranspiration} in the city of Chicago. Given that the evapotranspiration of the vegetation and crops is subject to seasonal changes in temperate zones such as Switzerland \cite{allen1998crop}, the values that correspond to the mid-season estimation (June to August) in \cite{nistor2016mapping}.
    % The albedo values are based on the work of Steward et al. \cite{stewart2012local}.
    % The source comma-separated value (CSV) file used in the computational workflow is available at \url{https://github.com/martibosch/lausanne-heat-islands/blob/master/data/raw/biophysical-table.csv}.
  \end{adjustwidth}
\end{table}

% \section{Relationships between the spatial patterns of tree canopy and air temperature}
% \label{sec:spatial-relationships}

% Exploration of relationships between the spatial patterns of tree canopy and air temperature, as Jupyter Notebook (IPYNB) \url{https://github.com/martibosch/lausanne-greening-scenarios/blob/master/notebooks/spatial-relationships.ipynb}.

\subsection*{Monitoring stations}
\label{sec:monitoring-stations}

The locations of the monitoring stations used to get the $T_{ref}$ and $UHI_{max}$ parameters of the InVEST urban cooling model are shown in \autoref{fig:monitoring-stations}. The operators of the stations are: Agrometeo, Federal roads office (ASTRA), Federal office for the environment (BAFU), General directorate for the environment of the Canton of Vaud (DGE), and the Federal Institute of Forest, Snow and Landscape Research (WSL) \cite{rebetez2018meteorological}.
The source CSV file with the operator, location and elevation in meters above sea level of the monitoring stations used in the computational workflow is available at \url{https://github.com/martibosch/lausanne-greening-scenarios/blob/master/data/raw/tair-stations/station-locations.csv}.
The code to produce \autoref{fig:monitoring-stations} is available as a Jupyter Notebook (IPYNB) at \url{https://github.com/martibosch/lausanne-greening-scenarios/blob/master/notebooks/stations.ipynb}.

\begin{figure}[ht]
  \centering
  % \includegraphics[width=.98\textwidth]{figures/monitoring-stations}
  \caption{\label{fig:monitoring-stations} Locations of the monitoring stations used to get the $T_{ref}$ and $UHI_{max}$ parameters. The basemap tile is provided by StamenDesign, under CC BY 3.0, with data from OpenStreetMap, under ODbL.}
\end{figure}

\section*{Results}

\subsection*{Effect of the proportion of tree cover}
\label{sec:effect-prop}

% The code for the evaluation of scenarios generated to explore the effect of increasing the proportion of tree cover
The code to produce the figures \ref{fig:greening-lulc-change}, \ref{fig:scenarios-prop}, \ref{fig:scenarios-prop-regplot} and \ref{fig:scenarios-prop-hists} is available as a Jupyter Notebook (IPYNB) at \url{https://github.com/martibosch/lausanne-greening-scenarios/blob/master/notebooks/scenarios-prop.ipynb}.

\begin{figure}[ht]
  \centering
  \includegraphics[width=.6\textwidth]{figures/scenarios-prop-regplot}
  \caption{\label{fig:scenarios-prop-regplot} Relationship between the proportion of candidate pixels transformed and the average simulated temperature $\overline{T}$ for each scenario sample. The translucent bands around the regression line represent the 95\% confidence intervals estimated using a bootstrap.}
\end{figure}

\begin{figure}[ht]
  \centering
  \includegraphics[width=.6\textwidth]{figures/scenarios-prop-hists}
  \caption{\label{fig:scenarios-prop-hists} Histogram of raster temperature values for a 25, 50, 75 and 100\% of the candidate pixels transformed. The temperature rasters for each histogram are computed by averaging the 10 simulations with the same proportion of candidate pixels transformed.}
\end{figure}

\subsection*{Effect of the spatial configuration of tree cover}
\label{sec:effect-config}

% Evaluation of scenarios generated to explore the effect of the spatial configuration of tree cover, as Jupyter Notebook (IPYNB) \url{https://github.com/martibosch/lausanne-greening-scenarios/blob/master/notebooks/scenarios-config.ipynb}.
The code to produce \autoref{fig:scenarios-config} is available as a Jupyter Notebook (IPYNB) at \url{https://github.com/martibosch/lausanne-greening-scenarios/blob/master/notebooks/scenarios-config.ipynb}.

\subsection*{Relationship between human exposure to higher air temperatures and the proportion of tree cover}
\label{sec:human-exposure}

% Evaluation of scenarios generated to explore the effect of the spatial configuration of tree cover, as Jupyter Notebook (IPYNB) \url{https://github.com/martibosch/lausanne-greening-scenarios/blob/master/notebooks/human-exposure.ipynb}.

The code to produce \autoref{fig:human-exposure} is available as a Jupyter Notebook (IPYNB) at \url{https://github.com/martibosch/lausanne-greening-scenarios/blob/master/notebooks/human-exposure.ipynb}.

\section*{Acknowledgments}
This research has been supported by the \'Ecole Polytechnique F\'ed\'erale de Lausanne (EPFL).

\nolinenumbers

% This is where your bibliography is generated. Make sure that your .bib file is actually called library.bib
\bibliography{references}

%This defines the bibliographies style. Search online for a list of available styles.
\bibliographystyle{unsrt}

\end{document}